---
title: "factorial anova ch13"
author: '2610922'
date: '2022-03-17'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r t1}
library(pwr)
library(rcompanion)
library(lsr)
library(car)
library(broom)
library(afex)
library(emmeans)
library(tidyverse)
library(performance)

factorial <- read_csv("Zhang et al. 2014 Exp 3.csv")%>%
  select(Condition, T1_Predicted_Interest_Composite, T2_Actual_Interest_Composite)%>%
  mutate(subject = row_number())%>%
  pivot_longer(names_to = "time",values_to = "interest", cols =       c("T1_Predicted_Interest_Composite","T2_Actual_Interest_Composite"))%>%
  mutate(Condition = dplyr::recode(Condition, "1" = "Ordinary", "2" = "Extraordinary"))%>%
  mutate(time = dplyr::recode(time, "T1_Predicted_Interest_Composite" = "time1_interest",
                       "T2_Actual_Interest_Composite" = "time2_interest")) %>%
  mutate(Condition = as.factor(Condition)) %>% 
  mutate (time= as.factor(time))
```

```{r t2}
sum_dat_factorial <- factorial %>%
  group_by(Condition, time) %>%
  summarise(mean=mean(interest, na.rm=TRUE),
            sd=sd(interest, na.rm=TRUE))
```

```{r t3}
ggplot(factorial, aes(x=time, y=interest, fill=Condition))+
  geom_violin(trim=FALSE, alpha=.4)+
  geom_boxplot(position=position_dodge(.9),
               width=.2,
               alpha=.6)+
  scale_x_discrete(labels=c("Time 1", "Time 2"))+
  stat_summary(fun="mean", geom="point",
               position = position_dodge(.9), width=.1)+
  stat_summary(fun.data="mean_se", geom="errorbar", width=.1, position=position_dodge(width=.9))+
  theme_minimal()
```

```{r t4}
ggplot(sum_dat_factorial, aes(x=time,y=mean, group=Condition, shape=Condition))+
  geom_point(size=3)+
  geom_line(aes(linetype=Condition))+
  scale_x_discrete(labels=c("Time 1","Time 2"))+
  theme_classic()
```

```{r t5}
mod_factorial <- aov_ez(id = "subject",
               data = factorial, 
               between = "Condition", 
               within = "time",
               dv = "interest", 
               type = 3) 
factorial_output <- anova(mod_factorial) %>% tidy()
```

```{r t6}
#normality
qqPlot(mod_factorial$lm$residuals)
shapiro.test(mod_factorial$lm$residuals)
#levene
test_levene(mod_factorial)
```

```{r t7}
# run the tests
posthoc_factorial <- emmeans(mod_factorial, 
                             pairwise ~ time| Condition, 
                             adjust = "bonferroni")
# tidy up the output of the tests
contrasts_factorial <- posthoc_factorial$contrasts %>%
  tidy()

posthoc_factorial2 <- emmeans(mod_factorial, 
                             pairwise ~ Condition| time, 
                             adjust = "bonferroni") 
contrasts_factorial2 <- posthoc_factorial2$contrasts %>%
  tidy()
#run effect size
d_extra_t1_t2 <- cohensD(interest ~ time, 
                         data = (filter(factorial, Condition == "Extraordinary") %>% droplevels())) 
d_ord_t1_t2 <- cohensD(interest ~ time, 
                         data = (filter(factorial, Condition == "Ordinary") %>% droplevels())) 
Condition_ds <- c(d_extra_t1_t2, d_ord_t1_t2)
contrasts_factorial <- contrasts_factorial %>%
  mutate(eff_size = Condition_ds)
d_time1_extra_ord <- cohensD(interest ~ Condition, 
                         data = (filter(factorial, time == "time1_interest") %>% droplevels())) 
d_time2_extra_ord <- cohensD(interest ~ Condition, 
                         data = (filter(factorial, time == "time2_interest") %>% droplevels()))
time_ds <- c(d_time1_extra_ord, d_time2_extra_ord)
contrasts_factorial2 <- contrasts_factorial2 %>%
  mutate(eff_size = time_ds)
```

```{r t8}
time_descrip <- factorial %>% 
  group_by(time) %>%
  summarise(mean_interest = mean(interest, na.rm = TRUE),
            sd_interest = sd(interest, na.rm = TRUE),
            min = mean(interest) - qnorm(0.975)*sd(interest)/sqrt(n()),
            max = mean(interest) + qnorm(0.975)*sd(interest)/sqrt(n()))
```
We conducted the same repeated measures ANOVA with interest as the dependent measure and again found a main effect of time, F(`r factorial_output$num.Df[2]`, `r factorial_output$den.Df[2]`) = `r factorial_output$statistic[2] %>% round(2)`, p < .001, ηp2 = `r factorial_output$ges[2] %>% round(3)`; anticipated interest at Time 1 (M = `r time_descrip$mean_interest[1] %>% round(2)`), SD = `r time_descrip$sd_interest[1]%>% round(2)`)) was lower than actual interest at Time 2 (M = `r time_descrip$mean_interest[2]%>% round(2)`, SD = `r time_descrip$sd_interest[2]%>% round(2)`).We also observed an interaction between time and type of experience, F(`r factorial_output$num.Df[3]`, `r factorial_output$den.Df[3]`) = `r factorial_output$statistic[3] %>% round(3)`, p = `r factorial_output$p.value[3] %>% round(2)`, ηp2 = `r factorial_output$ges[3] %>% round(3)`. Pairwise comparisons revealed that for ordinary events, predicted interest at Time 1 (M = `r sum_dat_factorial$mean[3]%>% round(2)`, SD = `r sum_dat_factorial$sd[3]%>% round(2)`) was lower than experienced interest at Time 2 (M = `r sum_dat_factorial$mean[4]%>% round(2)`, SD = `r sum_dat_factorial$sd[4]%>% round(2)`), t(`r contrasts_factorial$df[2]%>% round(2)`) = `r contrasts_factorial$statistic[2]%>% round(2)`, p < .001, d = `r contrasts_factorial$eff_size[2]%>% round(2)`. Although predicted interest for extraordinary events at Time 1 (M = `r sum_dat_factorial$mean[1]%>% round(2)`, SD = `r sum_dat_factorial$sd[1]%>% round(2)`) was lower than experienced interest at Time 2 (M = `r sum_dat_factorial$mean[2]%>% round(2)`, SD = `r sum_dat_factorial$sd[2]%>% round(2)`), t(`r contrasts_factorial$df[1]%>% round(2)`) = `r contrasts_factorial$statistic[1]%>% round(2)`, p < .001, d = `r contrasts_factorial$eff_size[1]%>% round(2)` , the magnitude of underestimation was smaller than for ordinary events.

```{r t9}
factorial <- factorial %>%
  mutate(interest_tukey = transformTukey(interest, plotit=TRUE))
tukey_factorial <- aov_ez(id = "subject",
               data = factorial, 
               between = "Condition", 
               within = "time",
               dv = "interest_tukey", 
               type = 3)
tukey_factorial
```
